A thermodynamic view of dusty protoplanetary disks: axisymmetric stability

### 2016-10-18 appendix.tex
notation change (k to kx), fig update, SI in main text

The explicit forms for $\dd\bm{F}$ and $\nabla\cdot\dd\bm{F}$, in the
radially-local approximation, are
\begin{align}
  &\dd F_r = - W F_r - \ii k Q,\\
  &\dd F_z = - W F_z - \left[Q^\prime + Q
    \left(\ln{\rho}\right)^\prime\right]  = - \ii\sigma \dd v_z, 
\end{align} 
where the last equality is the linearized
vertical momentum equation; and 
\begin{align}
  \nabla\cdot\dd\bm{F} &= F_r\left(\p_r\ln{\rho} - \ii k \right)W - F_z
  W^\prime - W \nabla\cdot\bm{F} + \ii kQ \p_r\ln{\rho} +
  \left(\ln{\rho}\right)^\prime\left[Q^\prime + Q
    \left(\ln{\rho}\right)^\prime\right] \notag\\
  & \phantom{=} + k^2 Q - \left\{Q^{\prime\prime} + 2
    Q^\prime\left(\ln{\rho}\right)^\prime
    + Q\left[\left(\ln{\rho}\right)^{\prime\prime}+\left(\ln{\rho}\right)^{\prime
      2}\right]\right\}\notag\\
    &= \left[\left(\p_r\ln{\rho}-\ii k\right)F_r - \nabla\cdot\bm{F} +
      F_z^\prime\right]W + \left(\ii k \p_r\ln{\rho} + k^2\right)Q  -
      \ii\sigma \dd v_z^\prime.
\end{align}

\section{Linearized dust diffusion}\label{lin_dust}
We consider small grains in the Epstein regime, with fixed internal
density and size, so that
\begin{align}
  \tstop  =  \frac{K}{\rho c_s}
\end{align}
\citep{price15}, where $K$ is a constant. Then the dust diffusion
function becomes
\begin{align}
  \mathcal{C} \equiv c_s^2\nabla\cdot\left(\tepsilon\tstop\nabla
  P\right) = - K c_s^2 \nabla\cdot
  \left(\frac{\tepsilon}{c_s}\bm{F}\right) =
  -Kc_s \left(\bm{F}\cdot\nabla\tepsilon + \tepsilon \nabla\cdot\bm{F}
  - \frac{1}{2}\tepsilon\bm{F}\cdot\nabla\ln{c_s^2}\right).  
\end{align}
Linearizing,
\begin{align}
  - \frac{\dd\mathcal{C}}{Kc_s} = \bm{F}\cdot\nabla\dd\tepsilon +
  \dd\bm{F}\cdot\nabla\tepsilon + \dd\tepsilon\nabla\cdot\bm{F} +
  \tepsilon\nabla\cdot\dd\bm{F} -
  \frac{1}{2}\nabla\ln{c_s^2}\cdot\left(\bm{F}\dd\tepsilon + \tepsilon
  \dd\bm{F}\right). 
\end{align}
The linearized dust-fraction $\dd\tepsilon$ and its derivatives in the
radially-local approximation are given by
\begin{align}
  \dd\tepsilon      &= (1-\tepsilon)W - \frac{Q}{c_s^2}, \\
  \p_r\dd\tepsilon &= \left[(1-\tepsilon)\left(\ii k -
    \p_r\ln{\rho}\right) - \p_r\tepsilon\right]W + \left(\p_r\ln{c_s^2}
  + \p_r\ln{\rho} - \ii k \right)\frac{Q}{c_s^2},\\
  \dd\tepsilon^\prime &= (1-\tepsilon)W^\prime - \tepsilon^\prime W -
  \left(\frac{Q}{c_s^2}\right)^\prime. 
\end{align}

@@

\section{Compressible streaming instability}\label{compressible_streaming}
We consider an unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =
0$. The background $\p_rP/\rho$ and $\tepsilon$ are constants. 
We Fourier analyze in $r$ and $z$ so that $\p_z\to \ii k_z$ and
$\p_r\to \ii k_x$ when acting on perturbations, and denote $|\bm{k}|^2
\equiv k_x^2 + k_z^2$. We neglect 
background gradients when compared with that of perturbations. The linearized
equations are, after eliminating the azimuthal velocity:

\begin{align}
  \ii \sigma W &=\nabla\cdot \bm{\dd v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\\
    \sigma^2 \dd v_x &= \kappa^2 \dd v_x - \ii \sigma F_r W +
    k_x\sigma Q,\\
  -\ii\sigma\dd v_z &= -\ii k_zQ,\\
\ii \zeta \sigma  Q & = \frac{P}{\rho} \nabla \cdot \bm{\dd v}   -
  \frac{\dd \mathcal{C}}{\rho}. 
\end{align}
The artifical factor $\zeta = 1$  is inserted to keep track of the
left-hand-side of the energy equation. We note that the incompressible
condition used by \cite{jacquet11} is obtained by setting $\zeta$ to zero. 
The linearized dust diffusion
function (Appendix \ref{lin_dust}), 
under the above approximations, is 
\begin{align}
-\frac{\dd \mathcal{C}}{\rho}  = \tstop c_s^2 \left[\ii k_x F_r\left( 1-2\tepsilon\right)
  W + \tepsilon |\bm{k}|^2 Q\right]. 
\end{align}

>>>>

The explicit forms for $\dd\bm{F}$ and $\nabla\cdot\dd\bm{F}$, in the
radially-local approximation, are
\begin{align}
  &\dd F_r = - W F_r - \ii k_x Q,\\
  &\dd F_z = - W F_z - \left[Q^\prime + Q
    \left(\ln{\rho}\right)^\prime\right]  = - \ii\sigma \dd v_z, 
\end{align} 
where the last equality is the linearized
vertical momentum equation; and 
\begin{align}
  \nabla\cdot\dd\bm{F} &= F_r\left(\p_r\ln{\rho} - \ii k_x \right)W - F_z
  W^\prime - W \nabla\cdot\bm{F} + \ii k_x Q \p_r\ln{\rho} +
  \left(\ln{\rho}\right)^\prime\left[Q^\prime + Q
    \left(\ln{\rho}\right)^\prime\right] \notag\\
  & \phantom{=} + k_x^2 Q - \left\{Q^{\prime\prime} + 2
    Q^\prime\left(\ln{\rho}\right)^\prime
    + Q\left[\left(\ln{\rho}\right)^{\prime\prime}+\left(\ln{\rho}\right)^{\prime
      2}\right]\right\}\notag\\
    &= \left[\left(\p_r\ln{\rho}-\ii k_x\right)F_r - \nabla\cdot\bm{F} +
      F_z^\prime\right]W + \left(\ii k_x \p_r\ln{\rho} + k_x^2\right)Q  -
      \ii\sigma \dd v_z^\prime.
\end{align}

\section{Linearized dust diffusion}\label{lin_dust}
We consider small grains in the Epstein regime, with fixed internal
density and size, so that
\begin{align}
  \tstop  =  \frac{K}{\rho c_s}
\end{align}
\citep{price15}, where $K$ is a constant. Then the dust diffusion
function becomes
\begin{align}
  \mathcal{C} \equiv c_s^2\nabla\cdot\left(\tepsilon\tstop\nabla
  P\right) = - K c_s^2 \nabla\cdot
  \left(\frac{\tepsilon}{c_s}\bm{F}\right) =
  -Kc_s \left(\bm{F}\cdot\nabla\tepsilon + \tepsilon \nabla\cdot\bm{F}
  - \frac{1}{2}\tepsilon\bm{F}\cdot\nabla\ln{c_s^2}\right).  
\end{align}
Linearizing,
\begin{align}
  - \frac{\dd\mathcal{C}}{Kc_s} = \bm{F}\cdot\nabla\dd\tepsilon +
  \dd\bm{F}\cdot\nabla\tepsilon + \dd\tepsilon\nabla\cdot\bm{F} +
  \tepsilon\nabla\cdot\dd\bm{F} -
  \frac{1}{2}\nabla\ln{c_s^2}\cdot\left(\bm{F}\dd\tepsilon + \tepsilon
  \dd\bm{F}\right). 
\end{align}
The linearized dust-fraction $\dd\tepsilon$ and its derivatives in the
radially-local approximation are given by
\begin{align}
  \dd\tepsilon      &= (1-\tepsilon)W - \frac{Q}{c_s^2}, \\
  \p_r\dd\tepsilon &= \left[(1-\tepsilon)\left(\ii k_x -
    \p_r\ln{\rho}\right) - \p_r\tepsilon\right]W + \left(\p_r\ln{c_s^2}
  + \p_r\ln{\rho} - \ii k_x \right)\frac{Q}{c_s^2},\\
  \dd\tepsilon^\prime &= (1-\tepsilon)W^\prime - \tepsilon^\prime W -
  \left(\frac{Q}{c_s^2}\right)^\prime. 
\end{align}

@@

\section{One-fluid dispersion relation for the streaming instability}\label{compressible_streaming}
We consider an unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =
0$. The background $\p_rP/\rho$ and $\tepsilon$ are constants. 
We Fourier analyze in $r$ and $z$ so that $\p_z\to \ii k_z$ and
$\p_r\to \ii k_x$ when acting on perturbations, and denote $|\bm{k}|^2
\equiv k_x^2 + k_z^2$. We neglect 
background gradients when compared with that of perturbations. The linearized
equations are, after eliminating the azimuthal velocity:

\begin{align}
  \ii \sigma W &=\nabla\cdot \dd \bm{ v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\\
    \sigma^2 \dd v_x &= \kappa^2 \dd v_x - \ii \sigma F_r W +
    k_x\sigma Q,\\
  -\ii\sigma\dd v_z &= -\ii k_zQ,\\
\ii \zeta \sigma  Q & = \frac{P}{\rho} \nabla \cdot \dd\bm{v}   -
  \frac{\dd \mathcal{C}}{\rho}. 
\end{align}
The artifical factor $\zeta = 1$  is inserted to keep track of the
left-hand-side of the energy equation. We note that the incompressible
condition used by \cite{jacquet11} is obtained by setting $\zeta$ to zero. 
The linearized dust diffusion
function (Appendix \ref{lin_dust}), 
under the above approximations, is 
\begin{align}
-\frac{\dd \mathcal{C}}{\rho}  = \tstop c_s^2 \left[\ii k_x F_r\left( 1-2\tepsilon\right)
  W + \tepsilon |\bm{k}|^2 Q\right]. 
\end{align}

### 2016-10-18 linear.tex
notation change (k to kx), fig update, SI in main text

\section{Linearized equations}\label{linear_problem}
To compute explicit solutions to the linear problem, 
we consider radially-localized axisymmetric disturbances of the form  
\begin{align}
  \delta f (r, z) = \delta f_1(r,z)\exp{(\ii k r)},
\end{align} 
%and similarly for $\dd P$ and $\dd\bm{v}$. 
where $k$ is a real wavenumber such that $|kr|\gg 1$, and the
amplitude $\dd f_1(r,z)$ is 
a slowly-varying function of $r$. Then 
$\p_r\to i k$ when acting on the above primitive perturbations, and we may
neglect curvature terms. We take  
$k>0$ without loss of generality. Hereafter, we drop the subscript 1
on the amplitudes. 
%The frequency $\sigma = \omega +
%\ii s$ is generally complex, with $\omega$ being the real frequency
%and $s$ is the real growth rate.

@@

\begin{align}
  \ii\sigma W &= \ii k \dd v_r + \dd v_z^\prime +
  \dd v_r \p_r\ln{\rho} + \dd v_z\p_z\ln{\rho},\label{lin_mass}\\
  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi 
% +  \delta\bm{F}\cdot\hat{\bm{r}}
- W F_r - \ii k Q,\label{lin_xmom}\\
  \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
    v_\phi}{\p z}\dd v_z, \label{lin_ymom}\\
  -\ii\sigma\dd v_z &= - W F_z - \left[Q^\prime + Q
    \left(\ln{\rho}\right)^\prime\right]  %\delta\bm{F}\cdot\hat{\bm{z}} 
,\label{lin_zmom}\\
  \ii\sigma Q &= \frac{P}{\rho}\left(\ii k \dd v_r + \dd
               v_z^\prime\right) + \frac{1}{\rho}\left(\dd v_r\p_rP + \dd v_z \p_zP\right)\notag\\
                &\phantom{=}-\frac{P}{\rho} \dd v_r\p_r
               \ln{c_s^2} %, \label{lin_energy} 
%+ \dd v_z \p_z\ln{c_s^2}\right),\label{lin_energy} 
               - \frac{\dd\mathcal{C}}{\rho},\label{lin_energy} 
\end{align}  
where $^\prime \equiv \p_z$ and recall $\bm{F} \equiv -\nabla P/\rho$. 
The linearized dust-duffusion function 
$\dd\mathcal{C}$ is given in  Appendix \ref{lin_dust}. 
% and $\dd\bm{F}$ is the linearized pressure
%force, given in Appendix \ref{lin_press}. 
Note that we have assumed a temperature profile that only depends on $r$.

@@

Eq. \ref{lin_mass}---\ref{lin_energy} is a set of ordinary
differntial equations in $z$. All coefficients and amplitudes are
evaluated at a fiducial radius $r=r_0$, but their full $z$-dependence
is retained. Numerical solutions are generally required for the 
verticall-stratified problem, as we pursue below.

However, in the special case of an unstratified disk (neglecting the vertical
component of the stellar gravity), we may also Fourier analyze in $z$
to obtain a local dispersion relation. This is done in 
Appendix \ref{compressible_streaming}. There we show the above linearized
equations give the streaming instability in the appropriate limit 
\citep{youdin05a,jacquet11}.

%\subsection{Boundary conditions}

%Eq. \ref{lin_mass}---\ref{lin_energy} can be reduced to a set of
%first-order differential equations for $W, Q$ and $\dd v_z$. We can
%see this schematically as follows. Eq. \ref{lin_xmom} and
%\ref{lin_ymom} may be combined to yield 
%\begin{align*}
%  \dd v_r = \dd v_r (W, Q, \dd v_z). 
%\end{align*} 
%We can then take the continuity equation as an equation for $\dd v_z$, 
%\begin{align*}
%\text{Eq. \ref{lin_mass}} \Rightarrow \dd v_z^\prime(z) = \dd
%  v_z^\prime(W,Q,\dd v_z),
%\end{align*}
%and the vertical momentum equation as an equation for $Q$, 
%\begin{align*}
%\text{Eq. \ref{lin_zmom}} \Rightarrow Q^\prime(z) = 
% Q^\prime(W,Q,\dd v_z). 
%\end{align*}

>>>>

\section{Linearized equations}\label{linear_problem}
To compute explicit solutions to the linear problem, 
we consider radially-localized axisymmetric disturbances of the form  
\begin{align}
  \delta f (r, z) = \delta f_1(r,z)\exp{(\ii k_x r)},
\end{align} 
%and similarly for $\dd P$ and $\dd\bm{v}$. 
where $k_x$ is a real wavenumber such that $|k_xr|\gg 1$, and the 
amplitude $\dd f_1(r,z)$ is 
a slowly-varying function of $r$. Then 
$\p_r\to i k_x$ when acting on the above primitive perturbations, and we may
neglect curvature terms. We take  
$k_x>0$ without loss of generality. Hereafter, we drop the subscript 1
on the amplitudes. 
%The frequency $\sigma = \omega +
%\ii s$ is generally complex, with $\omega$ being the real frequency
%and $s$ is the real growth rate.

@@

\begin{align}
  \ii\sigma W &= \ii k_x \dd v_r + \dd v_z^\prime +
  \dd v_r \p_r\ln{\rho} + \dd v_z\p_z\ln{\rho},\label{lin_mass}\\
  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi 
% +  \delta\bm{F}\cdot\hat{\bm{r}}
- W F_r - \ii k_x Q,\label{lin_xmom}\\
  \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
    v_\phi}{\p z}\dd v_z, \label{lin_ymom}\\
  -\ii\sigma\dd v_z &= - W F_z - \left[Q^\prime + Q
    \left(\ln{\rho}\right)^\prime\right]  %\delta\bm{F}\cdot\hat{\bm{z}} 
,\label{lin_zmom}\\
  \ii\sigma Q &= \frac{P}{\rho}\left(\ii k_x \dd v_r + \dd
               v_z^\prime\right) + \frac{1}{\rho}\left(\dd v_r\p_rP + \dd v_z \p_zP\right)\notag\\
                &\phantom{=}-\frac{P}{\rho} \dd v_r\p_r
               \ln{c_s^2} %, \label{lin_energy} 
%+ \dd v_z \p_z\ln{c_s^2}\right),\label{lin_energy} 
               - \frac{\dd\mathcal{C}}{\rho},\label{lin_energy} 
\end{align}  
where $^\prime \equiv \p_z$ and recall $\bm{F} \equiv -\nabla P/\rho$. 
The linearized dust-duffusion function 
$\dd\mathcal{C}$ is given in  Appendix \ref{lin_dust}. 
% and $\dd\bm{F}$ is the linearized pressure
%force, given in Appendix \ref{lin_press}. 
Note that we have assumed a temperature profile that only depends on $r$.

@@

Eq. \ref{lin_mass}---\ref{lin_energy} is a set of ordinary
differntial equations in $z$. All coefficients and amplitudes are
evaluated at a fiducial radius $r=r_0$, but their full $z$-dependence
is retained. We now discuss explicit solutions to these equations. 
% Numerical solutions are generally required for the 
%vertically-stratified problem.

\section{Compressible streaming instability} 
In the special case of an unstratified disk (neglecting the vertical
component of the stellar gravity), we may also Fourier analyze in $z$
to obtain a local dispersion relation of the form  
$\sum_{j=1}^{5}c_j(k_x,k_z)\sigma^j = 0$, where $k_z$ is a real
vertical wavenumber. The coefficients $c_j$ can be read 
off Eq. \ref{streaming_dispersion} in  Appendix \ref{compressible_streaming}. 
There we also show that this dispersion relation reduces to that for
the streaming instability in the limit of incompressible gas and small
$\tstop$ \citep{youdin05a,jacquet11}.

We solve Eq. \ref{streaming_dispersion} for selected cases where the  
growth rates of the streaming instability have been 
calculated analytically from full two-fluid disk models, as well as
numerically from particle-gas simulations 
\citep[namely][]{youdin07,bai10b}. These authors use normalized
wavenumbers $K_x = \eta r k_x$ and similarly for $K_z$, where
\begin{align} 
  \eta \equiv -\frac{1}{2\rhog r\OmK^2}\frac{\p P}{\p r} = 
  \frac{1}{2\left(1-\tepsilon\right)}\frac{F_r}{r\OmK^2}, 
\end{align} 
measures the pressure offset of Keplerian rotation. We use
$\eta=0.05c_s/r\OmK$. They also parametrize the stopping time with
$\tau_\mathrm{s}=\tstop/(1-\tepsilon)$.

Table \ref{si_compare} compares the eigenfrequencies obtained from the
one-fluid dispersion relation to that from the above studies. As
expected eigenfrequencies agree better with decreasing $\tstop$ since
in that limit the mixture behaves more like a single fluid. Most
importantly, we find the work done $\mathcal{W}>0$ in all cases, and
hence find growing oscillations. Thus the streaming instability stems
from the pressure evolution lagging behind that of the total density.

\begin{deluxetable*}{llrrrrr}
  \tablecolumns{7}
  \tablecaption{Selected eigenfrequencies for the streaming
    instability. \label{si_compare}
  }
  \tablehead{
    \colhead{Mode} & 
    \colhead{$\tau_\mathrm{s}\OmK$} &
    \colhead{$\epsilon$} &
     \colhead{$K_x = K_z$} &
      \colhead{$\sigma/\OmK$ (two-fluid)} &
    \colhead{$\sigma/\OmK$ (one-fluid, Eq. \ref{streaming_dispersion})}   &
    \colhead{$\mathcal{W}$ (arbitrary units) } 
  }
\startdata
 linA, \cite{youdin07} &  $0.1$       & 3.0 & 30    & $-0.3480 +
 0.4190\ii$ & $0.3647 + 0.4251\ii$ & $0.026$\\

linB, \cite{youdin07} & $0.1$        &  0.2 & 6 & $0.4999 +
0.0154\ii$&   $-0.4999 + 0.0054\ii$  & $1.74$ \\

linC,  \cite{bai10b}  & $10^{-2}$   &  2.0 & 1500&   $0.1049 +
0.5981\ii$   &  $0.1344 + 0.6665\ii$  & $3.58\times10^{-4}$\\

linD, \cite{bai10b} &  $10^{-3}$   &  2.0 & 2000 & $0.3225 +
0.3154\ii$& $0.3228 + 0.3155\ii$ &  $1.84\times10^{-4}$
\enddata
\end{deluxetable*}

%\subsection{Boundary conditions}
%Eq. \ref{lin_mass}---\ref{lin_energy} can be reduced to a set of
%first-order differential equations for $W, Q$ and $\dd v_z$. We can
%see this schematically as follows. Eq. \ref{lin_xmom} and
%\ref{lin_ymom} may be combined to yield 
%\begin{align*}
%  \dd v_r = \dd v_r (W, Q, \dd v_z). 
%\end{align*} 
%We can then take the continuity equation as an equation for $\dd v_z$, 
%\begin{align*}
%\text{Eq. \ref{lin_mass}} \Rightarrow \dd v_z^\prime(z) = \dd
%  v_z^\prime(W,Q,\dd v_z),
%\end{align*}
%and the vertical momentum equation as an equation for $Q$, 
%\begin{align*}
%\text{Eq. \ref{lin_zmom}} \Rightarrow Q^\prime(z) = 
% Q^\prime(W,Q,\dd v_z). 
%\end{align*}

### 2016-10-18 result.tex
notation change (k to kx), fig update, SI in main text

Fig. \ref{vsi_dust_loading} show unstable modes for different values
of $\epsilon_0$ with fixed perturbation wavenumber  $k\Hg = 30$. The
eigenvalue distributions for $\epsilon_0 \leq 10^{-2}$ are similar to the
dust-free fiducial case considered by \citetalias{lin15}. This is
expected since $\epsilon_0 < \hgas$ (\S\ref{vsi_est}). Eigenvalues 
consists 
of the roughly horizontal `body modes' and the nearly-vertical
`surface modes'. The latter is due to the imposed vertical boundaries
\citep{barker15}.

@@

In Fig. \ref{vsi_dust_loading_vareps} we plot the growth rates as a
function of $\epsilon_0$ for different perturbation wavenumbers $k$. We
find that dust-loading stabilizes the VSI more effectively for shorter
wavelength perturbations. This is because for high wavenumbers the
dominant modes are surface modes, which are effectively stabilized by
dust-loading since bouyancy forces are larges near the disk
boundaries. The figure suggest that VSI becomes much less efficient
for $\epsilon_0\gtrsim 0.1$ and $k\Hg\gtrsim 50$.

@@

Fig. \ref{result2d_fixZ} compares the fastest growing VSI body modes  
with $k\Hg=30$ for the two cases above.\citepalias[The thinner domain
  adopted here eliminates surface modes, ][]{lin15}.   
We find very similar mode 
frequencies 
\begin{align*}
  \sigma = \begin{cases}
    \left(0.3053\ii - 0.8142\right)h_\mathrm{g}\OmK & \Hd=0.99\Hg, \\
    \left(0.3178\ii - 1.2237\right)h_\mathrm{g}\OmK & \Hd=0.1\Hg,
  \end{cases}
\end{align*}
since the vertical shear profile is similar throughout most of the
disk. However, meridional motions are surpressed near the midplane of
the $\Hd=0.1\Hg$ disk, as expected from the larger buoyancy frequency
as compared to vertical shear there. This leads to a structure
analogous to PPD dead zones: a quiescent midplane between active
surface layers \citep{gammie96}.

@@

We shall consider ultra thin dust layers with $\Hd\leq 0.01\Hg$ and thus
restrict the vertical domain to $\zmax = 0.02\Hg$. We fix the 
metalicity $Z=0.01$ so the midpane dust-to-gas ratio $\epsilon_0$ is 
$O(1)$.  We consider radial wavenumbers such $k\Hd=1$.

@@

\subsection{Effect of a radially-varying dust-to-gas ratio}\label{varHd}
{\bf move to appendix?}

Here we allow the dust-to-gas ratio $\epsilon$ to depend on
$r$. Specifically we let $\epsilon_0\propto r^{-1}$ and
$H_\epsilon \propto \Hg$ (cf. constant values in the previous
calculations). Then
\begin{align*}
  \frac{\p\epsilon}{\p r} =
  \left(\frac{z^2}{H_\epsilon^2}\frac{d\ln{\Hg}}{dr} -
    \frac{1}{r}\right)\epsilon.   
\end{align*}
We fix $Z=0.01$, $\Hd=0.8\Hg$ at $r=r_0$, and
$k\Hg=30$.

>>>>

Fig. \ref{vsi_dust_loading} show unstable modes for different values
of $\epsilon_0$ with fixed perturbation wavenumber  $k_x\Hg = 30$. The
eigenvalue distributions for $\epsilon_0 \leq 10^{-2}$ are similar to the
dust-free fiducial case considered by \citetalias{lin15}. This is
expected since $\epsilon_0 < \hgas$ (\S\ref{vsi_est}). Eigenvalues 
consists 
of the roughly horizontal `body modes' and the nearly-vertical
`surface modes'. The latter is due to the imposed vertical boundaries
\citep{barker15}.

@@

In Fig. \ref{vsi_dust_loading_vareps} we plot the growth rates as a
function of $\epsilon_0$ for different perturbation wavenumbers $k$. We
find that dust-loading stabilizes the VSI more effectively for shorter
wavelength perturbations. This is because for high wavenumbers the
dominant modes are surface modes, which are effectively stabilized by
dust-loading since bouyancy forces are larges near the disk
boundaries. The figure suggest that VSI becomes much less efficient
for $\epsilon_0\gtrsim 0.1$ and $k_x\Hg\gtrsim 50$.

@@

Fig. \ref{result2d_fixZ} compares the fastest growing VSI body modes  
with $k_x\Hg=30$ for the two cases above.\citepalias[The thinner domain
  adopted here eliminates surface modes, ][]{lin15}.   
We find very similar mode 
frequencies 
\begin{align*}
  \sigma = \begin{cases}
    \left(0.3053\ii - 0.8142\right)h_\mathrm{g}\OmK & \Hd=0.99\Hg, \\
    \left(0.3178\ii - 1.2237\right)h_\mathrm{g}\OmK & \Hd=0.1\Hg,
  \end{cases}
\end{align*}
since the vertical shear profile is similar throughout most of the
disk. However, meridional motions are surpressed near the midplane of
the $\Hd=0.1\Hg$ disk, as expected from the larger buoyancy frequency
as compared to vertical shear there. This leads to a structure
analogous to PPD dead zones: a quiescent midplane between active
surface layers \citep{gammie96}.

@@

We shall consider ultra thin dust layers with $\Hd\leq 0.01\Hg$ and thus
restrict the vertical domain to $\zmax = 0.02\Hg$. We fix the 
metalicity $Z=0.01$ so the midpane dust-to-gas ratio $\epsilon_0$ is 
$O(1)$.  We consider radial wavenumbers such $k_x\Hd=1$.

@@

\subsection{Effect of a radially-varying dust-to-gas ratio}\label{varHd}
{\bf bonus stuff, maybe delete in final version}

Here we allow the dust-to-gas ratio $\epsilon$ to depend on
$r$. Specifically we let $\epsilon_0\propto r^{-1}$ and
$H_\epsilon \propto \Hg$ (cf. constant values in the previous
calculations). Then
\begin{align*}
  \frac{\p\epsilon}{\p r} =
  \left(\frac{z^2}{H_\epsilon^2}\frac{d\ln{\Hg}}{dr} -
    \frac{1}{r}\right)\epsilon.   
\end{align*}
We fix $Z=0.01$, $\Hd=0.8\Hg$ at $r=r_0$, and
$k_x\Hg=30$.
